---
title: "Analysis Storms and other severe weather eventsf in USA"
author: "Matias Cantergiani"
date: "5/10/2017"
output: html_document
---


## 1: Synopsis
The goal of the assignment is to explore the NOAA Storm Database and explore the effects of severe weather events on both population and economy.The database covers the time period between 1950 and November 2011.

The following analysis investigates which types of severe weather events are most harmful on:

1. Health (injuries and fatalities) 
2. Property and crops (economic consequences)

Information on the Data: [Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)

## 2: Data Processing

### 2.1: Data Loading

Download the raw data file and extract the data into a dataframe.Then convert to a data frame
```{r DataLoading}
library(readxl)
library(dplyr)
library(readr)
csv <- read.csv("C:/Users/mccan/Documents/R/Repositorios/ProgrammingAssignment5-2/repdata_data_StormData.csv", na.strings = "NA")
df <- as.data.frame(csv)
```

### 2.2: Examining Column Names and the classes

```{r ColumnNames}
colnames(df)
str(df)
```

### 2.4: Converting Exponent Columns into Actual Exponents instead of (M H, K, etc)

Making the PROPDMGEXP and CROPDMGEXP columns cleaner so they can be used to calculate property and crop cost.
```{r CorrectingExponents, results="hide"}
df <- df %>%
  mutate(PROPDMGEXP= as.numeric(df$PROPDMGEXP))

df$PROPDMGEXP[df$PROPDMGEXP == "H"] <- 2
df$PROPDMGEXP[df$PROPDMGEXP == "K"] <- 3
df$PROPDMGEXP[df$PROPDMGEXP == "M"] <- 6
df$PROPDMGEXP[df$PROPDMGEXP == "B"] <- 9

df <- df %>%  
  mutate(CROPDMGEXP= as.numeric(df$CROPDMGEXP))

df$CROPDMGEXP[df$CROPDMGEXP == "K"] <- 3   
df$CROPDMGEXP[df$CROPDMGEXP == "M"] <- 6   
df$CROPDMGEXP[df$CROPDMGEXP == "B"] <- 9   
```

### 2.5: Making Economic Cost Columns

```{r EconomicCostColumns}
df <- df %>%
  mutate(CostoProp = df$PROPDMG*(10^df$PROPDMGEXP)) %>%
  mutate(CostCrop = df$CROPDMG*(10^df$CROPDMGEXP))
```


## 3: Results

### 3.1: Events that are Most Harmful to Population Health

```{r HealthResults}
df_agrupado <- df %>%
  group_by(EVTYPE) %>%
  summarise(MeanFATALITIES = mean(FATALITIES), MeanINJURIES = mean(INJURIES) , Count = n()) %>%
  mutate(TotalDMG = MeanFATALITIES*0.75 + MeanINJURIES*0.25) %>%
  arrange( desc(TotalDMG), by_group = TRUE)

df_agrupado  
```

### 3.2: Events that have the Greatest Economic Consequences

```{r EconConsequences}
df_agrupado2 <- df %>%
  group_by(EVTYPE) %>%
  summarise(MeanPropCost = mean(CostoProp), MeanCropCost = mean(CostCrop) , Count = n()) %>%
  mutate(TotalCost = MeanPropCost + MeanCropCost) %>%
  arrange( desc(TotalCost), by_group = TRUE)

df_agrupado2 
```

